---
title: "Premier League Forwards 2022-23"
author: "Parthiv Purkayastha (github.com/pparthiv)"
date: "2022-09-30"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook presents the statistics of current Premier League 22/23 Season Forwards acquired through [FBRef](https://fbref.com/en/) (powered by [StatsBomb](https://statsbomb.com/)).

The notebook embeds an R shiny app which consists of options to compare players through visualizations or to rank them among others through the help of visualizations.

Unfortunately, some players are unavailable to be selected as they do not fulfill the [criteria](https://fbref.com/en/about/scouting-reports-explained) selected.

```{r, include=FALSE}
library(tidyverse)
library(dplyr)
library(fmsb)
library(shiny)
library(shinythemes)
library(shinydashboard)
library(DT)
library(ggplot2)
library(ggrepel)
library(plotly)
```

```{r, include=FALSE}
path <- 'Leagues/Premier League/prem_fw_py.csv'
prem <- read.csv(path)
prem <- prem[,-1]
rownames(prem) <- prem[,1]
prem <- prem[,-1]

tmp <- data.frame(prem[21,])

prem <- mutate_all(prem, function(x) as.numeric(as.character(x)))

for (i in rownames(prem[-21,])) {
  prem[i,'Max'] <- max(prem[c(i),2:98])
}

for (i in rownames(prem[-21,])) {
  prem[i,'Min'] <- min(prem[c(i),2:98])
}

prem['Team',] <- tmp
```

```{r, include=FALSE}
radar_chart <- function(name1, name2, name3, name4, name5){
  st1 <- gsub("\\ ", ".", name1)
  st2 <- gsub("\\ ", ".", name2)
  st3 <- gsub("\\ ", ".", name3)
  st4 <- gsub("\\ ", ".", name4)
  st5 <- gsub("\\ ", ".", name5)
  
  v_label <- rownames(prem)
  
  df <- prem[,c('Max', 'Min', st1, st2, st3, st4, st5)]
  df <- t(df)
  df <- data.frame(df)
  df <- mutate_all(df, function(x) as.numeric(as.character(x)))
  
  colr <- c(rgb(0.8, 0, 0, 1),rgb(0, 0.8, 0.8, 1),rgb(1, 1, 0, 1),rgb(1, 0.4, 0, 1),rgb(0, 1, 0.2, 1))
  colr_f <- c(rgb(0.8, 0, 0, 0.5),rgb(0, 0.8, 0.8, 0.5),rgb(1, 1, 0, 0.5),rgb(1, 0.4, 0, 0.5),rgb(0, 1, 0.2, 0.5))
  
  radarchart(df, cglty = 1, pcol = colr, pfcol = colr_f, plwd = 2, plty = 1, vlabels = v_label, vlcex = 0.9)
}

scatter_fw <- function(value1, value2, df){
  df <- df[-c(99:100)]
  df <- t(df)
  df <- data.frame(df)
  df[,-21] <- mutate_all(df[,-21], function(x) as.numeric(as.character(x)))
  df$Name <- gsub("\\.", " ", rownames(df))
  p <- ggplot(df, aes_string(x = value1, y = value2, label = 'Name')) +
    geom_point(aes(color = Team)) +
    geom_smooth(method=lm , se=TRUE) +
    theme(panel.background = element_blank(), panel.grid.major = element_blank(),panel.grid.minor = element_blank(),axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA, size=0.5)) +
    scale_x_continuous(name = NULL) + scale_y_continuous(name = NULL)
  ggplotly(p, width = 800)
}

avg_comp <- function(value1){
  name <- gsub("\\ ", ".", value1)
  temp <- data.frame(prem[-21,])
  temp <- temp[-c(98:99)]
  temp <- mutate_all(temp, function(x) as.numeric(as.character(x)))
  for (i in rownames(temp)) {
    temp[c(i),] <- rank(temp[c(i),])
  }
  temp <- mutate_all(temp, function(x) as.numeric(as.character(x)))
  
  player <- data.frame(temp[,c(name)])
  player <- mutate_all(player, function(x) as.numeric(as.character(x)))
  colnames(player) <- c(name)
  player$Categories <- rownames(temp)

  p <- ggplot(player, aes_string(x=name, y='Categories')) +
    geom_segment(aes_string(x=50.0, xend=name, y='Categories', yend='Categories'), color = "#7aa9ff") +
    geom_point(color = "#faa039", size=3) +
    theme_light() +
    theme(panel.background = element_blank(),axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA, size=0.5)) +
    scale_x_continuous(name = value1)
  ggplotly(p, width = 800)
}


prem_op <- function(df){
  df <- df[-c(99:100)]
  df <- t(df)
  rownames(df) <- gsub("\\.", " ", rownames(df))
  df <- data.frame(df[-1,])
  DT::datatable(df, options = list(scrollX = TRUE))
}

players <- gsub("\\.", " ", colnames(prem))
```

## Percentile Rank

This plot presents a player's performance compared to the 50th percentile in the league in the shape of a horizontal line chart.

```{r, echo=FALSE}

inputPanel(
  selectInput(
            label = "Forwards: ",
            'stats',
            choices = players[-c(99:100)],
            )
)

renderPlotly({
    avg_comp(input$stats)
  })

```

## Radar Chart

The following chart presents the radar-chart (or a spider plot) of a player's performance among various categories. These categories have been selected by the 'SCOUTING REPORT' panel of FBRef.

```{r, echo=FALSE}
inputPanel(
  selectInput(
            label = 'Player Red',
            'player1',
            choices = players[-c(99:100)]
            ),
          selectInput(
            label = 'Player Blue',
            'player2',
            choices = players[-c(99:100)]
            ),
          selectInput(
            label = 'Player Yellow',
            'player3',
            choices = players[-c(99:100)]
            ),
          selectInput(
            label = 'Player Orange',
            'player4',
            choices = players[-c(99:100)]
            ),
          selectInput(
            label = 'Player Green',
            'player5',
            choices = players[-c(99:100)]
            )
)

renderPlot({
    radar_chart(input$player1, input$player2, input$player3, input$player4, input$player5)}, height = 750, width = 750)
```

## Scatter Plot

The following plot shows the performance among the categories and returns a relation among those two. The grey zone is the confidence level of the values.

```{r, echo=FALSE}

inputPanel(
  selectInput(
                  label = "Category (X-axis): ",
                  'cat1',
                  choices = c(
                    "Non-Penalty Goals" = "Non.Penalty.Goals",
                    "Non-Penalty xG" = "Non.Penalty.xG",
                    "Shots Total" = "Shots.Total",
                    "Assists" = "Assists",
                    "xG Assisted" = "xG.Assisted",
                    "Non-Penalty xG + xG Assisted" = "npxG...xA",
                    "Shot Creating Actions" = "Shot.Creating.Actions",
                    "Pass Completion (in %)" = "Pass.Completion",
                    "Progressive Passes" = "Progressive.Passes",
                    "Progressive Carries" = "Progressive.Carries",
                    "Dribbles Completed" = "Dribbles.Completed",
                    "Touches in Attacking Penalty Area" = "Touches..Att.Pen.",
                    "Progressive Passes Received" = "Progressive.Passes.Rec",
                    "Pressures" = "Pressures",
                    "Tackles" = "Tackles",
                    "Interceptions" = "Interceptions",
                    "Blocks" = "Blocks",
                    "Clearances" = "Clearances",
                    "Aerials Won" = "Aerials.won"
                  )
                ),
                selectInput(
                  label = "Category (Y-axis): ",
                  'cat2',
                  choices = c(
                    "Non-Penalty Goals" = "Non.Penalty.Goals",
                    "Non-Penalty xG" = "Non.Penalty.xG",
                    "Shots Total" = "Shots.Total",
                    "Assists" = "Assists",
                    "xG Assisted" = "xG.Assisted",
                    "Non-Penalty xG + xG Assisted" = "npxG...xA",
                    "Shot Creating Actions" = "Shot.Creating.Actions",
                    "Pass Completion (in %)" = "Pass.Completion",
                    "Progressive Passes" = "Progressive.Passes",
                    "Progressive Carries" = "Progressive.Carries",
                    "Dribbles Completed" = "Dribbles.Completed",
                    "Touches in Attacking Penalty Area" = "Touches..Att.Pen.",
                    "Progressive Passes Received" = "Progressive.Passes.Rec",
                    "Pressures" = "Pressures",
                    "Tackles" = "Tackles",
                    "Interceptions" = "Interceptions",
                    "Blocks" = "Blocks",
                    "Clearances" = "Clearances",
                    "Aerials Won" = "Aerials.won"
                  )
                )
)

renderPlotly({
    scatter_fw(input$cat1, input$cat2, prem)
  })
```
